Assessing primitives performance\\ on multi-stage execution

### 2018-05-09 paper.tex
fixed appendixes

\todo{Add some intro sentence here and there + maybe ASM in appendix}
%% Appendix
\appendix
\section{Smart Syntax code}
\label{SmartSyntaxAppendix}

@@

\cba{Source code et generated C code et generated ASM code}

\section{Slang Plugin code}
\label{PluginAppendix}

@@

\cba{Source code et generated C code et generated ASM code}

\section{Slang code}
\label{SlangAppendix}

@@

	self methodReturnInteger: result
\end{code}
\begin{code}
rawCompare: string1 length: strLength1 with: string2 length: strLength2 accessBlock: accessBlock
	| c1 c2 min |
	<inline: true> "needs to be forced else slang does not inline it by default"
	min := strLength1 min: strLength2.
	0 to: min-1 do: 
		[:i | c1 := accessBlock value: string1 value: i.
			c2 := accessBlock value: string2 value: i.
			c1 = c2 ifFalse: [^c1 - c2]].
	^strLength1 - strLength2
\end{code}

\cba{Source code et generated C code et generated ASM code}

\section{Cog's RTL code}
\label{RTLAppendix}

\begin{code}
genPrimitiveStringCompareWith
	| instr jump jumpAbove jumpIncorrectFormat1 jumpIncorrectFormat2 jumpIncorrectFormat3 jumpIncorrectFormat4 jumpMidFailure jumpSuccess minSizeReg string1CharOrByteSizeReg string2CharOrByteSizeReg string1Reg string2Reg |
	<var: #jumpIncorrectFormat1 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat2 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat3 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat4 type: #'AbstractInstruction *'>
	<var: #jumpAbove type: #'AbstractInstruction *'>
	<var: #jumpSuccess type: #'AbstractInstruction *'>
	<var: #jump type: #'AbstractInstruction *'>
	<var: #jumpMidFailure type: #'AbstractInstruction *'>
	"I redefine those name to ease program comprehension"
	string1Reg := ReceiverResultReg.
	string2Reg := Arg0Reg.
	string1CharOrByteSizeReg := Arg1Reg.
	string2CharOrByteSizeReg := ClassReg.
	minSizeReg := SendNumArgsReg.
	"Load arguments in reg"
	cogit genLoadArgAtDepth: 0 into: string2Reg.
	"checks if string1 is a byteobject and get its size in bytes"
	self genGetFormatOf: string1Reg into: TempReg.
	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
	jumpIncorrectFormat1 := cogit JumpLess: 0.
	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
	jumpIncorrectFormat2 := cogit JumpAboveOrEqual: 0.	
	self genGetNumSlotsOf: string1Reg into: string1CharOrByteSizeReg.
	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string1CharOrByteSizeReg). 
	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg. 
	cogit SubR: TempReg R: string1CharOrByteSizeReg. 
	"checks if string2 is a byteobject and get its size in bytes"
	self genGetFormatOf: string2Reg into: TempReg.
	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
	jumpIncorrectFormat3 := cogit JumpLess: 0.
	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
	jumpIncorrectFormat4 := cogit JumpAboveOrEqual: 0.	
	self genGetNumSlotsOf: string2Reg into: string2CharOrByteSizeReg.
	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string2CharOrByteSizeReg).
	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg.
	cogit SubR: TempReg R: string2CharOrByteSizeReg.
	"Type and number of arguments are correct"
	"Compute the min"	 
	cogit CmpR: string1CharOrByteSizeReg R: string2CharOrByteSizeReg.
	jumpAbove := cogit JumpBelow: 0. 
	cogit MoveR: string1CharOrByteSizeReg R: minSizeReg. 
	jump := cogit Jump: 0. 
	jumpAbove jmpTarget: (cogit MoveR: string2CharOrByteSizeReg R: minSizeReg). 
	jump jmpTarget: (cogit CmpCq: 0 R: minSizeReg). 
	jumpSuccess := cogit JumpZero: 0. "if one of the string is empty, no need to go through the comparing loop"
	"Compare the bytes"
	cogit MoveCq: objectMemory baseHeaderSize  R: TempReg.
	cogit AddCq: objectMemory baseHeaderSize R: minSizeReg.
	instr := cogit MoveXbr: TempReg R: string1Reg R: string1CharOrByteSizeReg.
	cogit MoveXbr: TempReg R: string2Reg R: string2CharOrByteSizeReg.
	cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg. 
	jumpMidFailure := cogit JumpNonZero: 0. "the 2 compared characters are different, exit the loop"
	cogit AddCq: 1 R: TempReg.
	cogit CmpR: TempReg R: minSizeReg. 
	cogit JumpNonZero: instr.
	"all bytes from 1 to minSize are equal"
	self genGetNumBytesOf: string1Reg into: string1CharOrByteSizeReg.
	self genGetNumBytesOf: string2Reg into: string2CharOrByteSizeReg.
	jumpSuccess jmpTarget: (cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg).
	jumpMidFailure  jmpTarget: (cogit MoveR: string1CharOrByteSizeReg R: ReceiverResultReg).	
	self genConvertIntegerToSmallIntegerInReg: ReceiverResultReg.
	cogit genPrimReturn.
	jumpIncorrectFormat4 
		jmpTarget: (jumpIncorrectFormat3 
			jmpTarget: (jumpIncorrectFormat2 
				jmpTarget: (jumpIncorrectFormat1 jmpTarget: cogit Label))).
	^ CompletePrimitive}
\end{code}

\cba{Source code et generated ASM code}

>>>>

%% Appendix
\vfill\null %switch column
\appendix
\section{Smart Syntax code}
\label{SmartSyntaxAppendix}

Source code of the smart syntax method, which is used to generate the C code for the plugin but also as Smalltalk fall-back code.

@@

\vfill\null %switch column
\section{Slang Plugin code}
\label{PluginAppendix}

Source code of the plugin method generating the C code for the Slang plugin code. Most operations needs to go through the interpreter proxy.

@@

\vfill\null %switch column
\section{Slang code}
\label{SlangAppendix}

Source code of the slang method generating the C code for the Slang string comparison primitive. We split it in two methods, and forced inlining at Slang-to-C compilation time to remove the closures while avoiding code duplication.

@@

	self methodReturnInteger: result

rawCompare: string1 length: strLength1 with: string2 length: strLength2 accessBlock: accessBlock
	| c1 c2 min |
	<inline: true> "needs to be forced else slang does not inline it by default"
	min := strLength1 min: strLength2.
	0 to: min-1 do: 
		[:i | c1 := accessBlock value: string1 value: i.
			c2 := accessBlock value: string2 value: i.
			c1 = c2 ifFalse: [^c1 - c2]].
	^strLength1 - strLength2
\end{code}

\vfill\null %switch column
\section{Cog's RTL code}
\label{RTLAppendix}

Source code of the method generating the native code of the string comparison primitive on top of Cog's RTL. Method names starting with uppercase (in blue) are directly generating an abstract assembly instruction. Other methods called, starting with gen, generate a sequence of abstract assembly instructions. We duplicated the code checking if each operand is a ByteString and extract its size to generate better native code than doing the two operations separatedly.

\begin{code}
genPrimitiveStringCompareWith
	| instr jump jumpAbove jumpIncorrectFormat1 jumpIncorrectFormat2 jumpIncorrectFormat3 jumpIncorrectFormat4 jumpMidFailure jumpSuccess minSizeReg string1CharOrByteSizeReg string2CharOrByteSizeReg string1Reg string2Reg |
	<var: #jumpIncorrectFormat1 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat2 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat3 type: #'AbstractInstruction *'>
	<var: #jumpIncorrectFormat4 type: #'AbstractInstruction *'>
	<var: #jumpAbove type: #'AbstractInstruction *'>
	<var: #jumpSuccess type: #'AbstractInstruction *'>
	<var: #jump type: #'AbstractInstruction *'>
	<var: #jumpMidFailure type: #'AbstractInstruction *'>
	"I redefine those name to ease program comprehension"
	string1Reg := ReceiverResultReg.
	string2Reg := Arg0Reg.
	string1CharOrByteSizeReg := Arg1Reg.
	string2CharOrByteSizeReg := ClassReg.
	minSizeReg := SendNumArgsReg.
	"Load arguments in reg"
	cogit genLoadArgAtDepth: 0 into: string2Reg.
	"checks if string1 is a byteobject and get its size in bytes"
	self genGetFormatOf: string1Reg into: TempReg.
	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
	jumpIncorrectFormat1 := cogit JumpLess: 0.
	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
	jumpIncorrectFormat2 := cogit JumpAboveOrEqual: 0.	
	self genGetNumSlotsOf: string1Reg into: string1CharOrByteSizeReg.
	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string1CharOrByteSizeReg). 
	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg. 
	cogit SubR: TempReg R: string1CharOrByteSizeReg. 
	"checks if string2 is a byteobject and get its size in bytes"
	self genGetFormatOf: string2Reg into: TempReg.
	cogit CmpCq: objectMemory firstByteFormat R: TempReg.
	jumpIncorrectFormat3 := cogit JumpLess: 0.
	cogit CmpCq: objectMemory firstCompiledMethodFormat R: TempReg.
	jumpIncorrectFormat4 := cogit JumpAboveOrEqual: 0.	
	self genGetNumSlotsOf: string2Reg into: string2CharOrByteSizeReg.
	(cogit LogicalShiftLeftCq: objectMemory shiftForWord R: string2CharOrByteSizeReg).
	cogit AndCq: objectMemory wordSize - 1 R: TempReg R: TempReg.
	cogit SubR: TempReg R: string2CharOrByteSizeReg.
	\end{code}
	\vfill\null %switch column
	\begin{code}
	"Type and number of arguments are correct"
	"Compute the min"	 
	cogit CmpR: string1CharOrByteSizeReg R: string2CharOrByteSizeReg.
	jumpAbove := cogit JumpBelow: 0. 
	cogit MoveR: string1CharOrByteSizeReg R: minSizeReg. 
	jump := cogit Jump: 0. 
	jumpAbove jmpTarget: (cogit MoveR: string2CharOrByteSizeReg R: minSizeReg). 
	jump jmpTarget: (cogit CmpCq: 0 R: minSizeReg). 
	jumpSuccess := cogit JumpZero: 0. "if one of the string is empty, no need to go through the comparing loop"
	"Compare the bytes"
	cogit MoveCq: objectMemory baseHeaderSize  R: TempReg.
	cogit AddCq: objectMemory baseHeaderSize R: minSizeReg.
	instr := cogit MoveXbr: TempReg R: string1Reg R: string1CharOrByteSizeReg.
	cogit MoveXbr: TempReg R: string2Reg R: string2CharOrByteSizeReg.
	cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg. 
	jumpMidFailure := cogit JumpNonZero: 0. "the 2 compared characters are different, exit the loop"
	cogit AddCq: 1 R: TempReg.
	cogit CmpR: TempReg R: minSizeReg. 
	cogit JumpNonZero: instr.
	"all bytes from 1 to minSize are equal"
	self genGetNumBytesOf: string1Reg into: string1CharOrByteSizeReg.
	self genGetNumBytesOf: string2Reg into: string2CharOrByteSizeReg.
	jumpSuccess jmpTarget: (cogit SubR: string2CharOrByteSizeReg R: string1CharOrByteSizeReg).
	jumpMidFailure  jmpTarget: (cogit MoveR: string1CharOrByteSizeReg R: ReceiverResultReg).	
	self genConvertIntegerToSmallIntegerInReg: ReceiverResultReg.
	cogit genPrimReturn.
	jumpIncorrectFormat4 
		jmpTarget: (jumpIncorrectFormat3 
			jmpTarget: (jumpIncorrectFormat2 
				jmpTarget: (jumpIncorrectFormat1 jmpTarget: cogit Label))).
	^ CompletePrimitive}
\end{code}
